---
title: "Visualisation in R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(tidyverse)
library(RColorBrewer)
library(learnr)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE)

economic_data <- read_csv("https://raw.githubusercontent.com/drros84/euro_cluster_app/master/macro_dataset.csv") %>% 
  select(country = region, year = Year, GDP, wages = average_annual_wage, inflation = cpi, 
         productivity, unemployment = unemployment_rate, research_investment = GERD, govt_debt) %>% 
  na.omit() 
```

## Introduction

This training will introduce you to data visualisation in R. By the end, you should be able to produce a range of charts, including adding some interactivity. Note that you can find more useful information on ggplot2 functions in the ggplot cheatsheet [here](https://www.rstudio.com/resources/cheatsheets/). 

**Note that this tutorial only covers the very basics of data visualisation using ggplot and there is a lot more to learn.**

We will be using a dataset containing economic data for European countries from 1995 to 2014. 

It covers data for the following countries:

```{r intro-example1, echo = TRUE}

unique(economic_data$country)

```

Now, use the same code to look at how many years are covered in the data:

```{r unique1, exercise=TRUE}

```

```{r unique1-solution}

unique(economic_data$year)

```

You can look at the dataset by using the ``head()`` function:

```{r head, exercise=TRUE}

```

```{r head-solution}

head(economic_data)

```


### ggplot2 

The plotting package from the tidyverse that we're going to use is **ggplot2**. It was developed by [Hadley Wickham](https://hadley.nz), who also led development of the rest of the **tidyverse** group of packages. This is useful because it makes the two very compatible. 

Additionally, **ggplot2** was developed on some solid data visualisation theory - in particular, it seeks to implement Wilkinson's [Grammar of Graphics](https://vita.had.co.nz/papers/layered-grammar.html) approach to data visualisation. This means that it will generally force you to approach sound data visualisation principles.

There are 3 key parts to a ggplot chart:

* **data**

* **aesthetics** - which scales to map the data onto

* **geometries** - how the plot will look

You build up a plot by adding elements/layers.

For example, in the chart below, using data for the top 10 countries with the highest GDP in 2014, and we only define the **data** and **aesthetics** - but not the **geometries**. As a result, all we see is an empty chart, but you can see the scale, with the corresponding axes, are already defined here:

```{r introexample, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = country, y = GDP))

```

## Geometries

In this section, we experiment with several geometries (shapes).

### Bar chart

You can create a bar chart of spending by agency in 2020 using ``geom_col()``. **Notice that for ggplot, we use + between each line rather than %>%**:

```{r geom_bar_ex1, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = country, y = GDP)) +
  geom_col()

```

Notice that the order is not particularly useful. We can easily reorder it by decreasing order of ``values`` using the ``reorder()`` function:

```{r geom_bar_ex2, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col()

```

Now, create a bar chart of ``wages`` per country, for the countries with top 10 highest wages in 2014. As in the chart above, order the bar chart in decreasing order of ``wages``:

```{r geom_col1, exercise=TRUE}

```

```{r geom_col1-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(wages)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -wages), y = wages)) +
  geom_col()

```

### Scatter plot

We can use the ``geom_point`` to create a scatterplot. Below, we plot ``inflation`` against ``unemployment``, with the colour for each country, in 2014:

```{r geom_point_ex1, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = unemployment, y = inflation)) +
  geom_point()

```

We can see there is definitely a negative correlation between inflation and unemployment - interesting!

To make that clear, we can also add a trendline to the chart:

```{r geom_point_ex2, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = unemployment, y = inflation)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

Additionally, we can change the size of the points is linked to the ``wages``:

```{r geom_point_ex3, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = unemployment, y = inflation, size = wages)) +
  geom_point() 

```

Now, create a scatterplot of ``productivity`` against ``research_investment``, where size is linked to ``wages``, for 2014. You do not need a trendline:

```{r geom_point1, exercise=TRUE}

```

```{r geom_point1-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = productivity, y = research_investment, size = wages)) +
  geom_point()

```


### Line chart

Next, we draw a line chart of GDP over time for France using the ``geom_line()`` function:

```{r geom_line_ex1, echo = TRUE}

economic_data %>% 
  filter(country == "France") %>% 
  ggplot(aes(x = year, y = GDP)) +
  geom_line()

```

Now, draw a line chart of ``wages`` over time for Italy:

```{r geom_line1, exercise=TRUE}

```

```{r geom_line1-solution}

economic_data %>% 
  filter(country == "Italy") %>% 
  ggplot(aes(x = year, y = wages)) +
  geom_line()

```

### Text

You can even replace points by text on a chart using ``geom_text``. This is the same chart as the scatterplot of unemployment and inflation, except that we put the country names instead of points:

```{r geom_text_ex1, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = unemployment, y = inflation, label = country)) +
  geom_text()

```


Draw a text chart of ``GDP`` against ``govt_debt`` in 2014, using ``geom_text()`` to display country names instead of points:

```{r geom_text1, exercise=TRUE}

```

```{r geom_text1-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = GDP, y = govt_debt, label = country)) +
  geom_text()

```


### Combining geometries

Note that you can of course combine several geometries in one chart. For example, this chart combines a bar chart of ``productivity`` with a scatter plot of ``govt_debt`` :

```{r geom_combined_ex1, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(productivity)) %>% 
  head(10) %>% 
  ggplot(aes(x = country)) +
  geom_col(aes(y = productivity)) +
  geom_point(aes(y = govt_debt))

```

## Colour

### Basics

Now, we look at how to colour charts. The most basic way of adding colour to a chart is very simple, like for this bar chart:

```{r col_ex1, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = unemployment, y = inflation)) +
  geom_point(col = "red")

```

Note that in the above chart, we used the ``col`` argument. For a bar chart, we will need to use ``fill``:

```{r col_ex2, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col(fill = "red")

```

In the examples above, the colour is the same for all points/bars. If you want the colour the represent a variable, you need to put it inside the ``aes()`` function, as follows:

```{r col_ex3, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = country)) +
  geom_point()

```


Notice that the chart above has too many colours to be useful. Instead, in the chart below, we limit the number of colours by recoding the country names so that we only show the names of the top 7 countries by unemployment, and the rest as "other":

```{r col_ex4, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(unemployment)) %>% 
  mutate(country = ifelse(row_number() <= 7, country, "Other")) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = country)) +
  geom_point()

```

Do you see how the legend shows "Other" in the middle of all the countries, because the legend is alphabetical? That is not very helpful. To change that, I recode "country" as a factor, which will be in decreasing order of ``unemployment``:

```{r col_ex5, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(unemployment)) %>% 
  mutate(country = ifelse(row_number() <= 7, country, "Other")) %>% 
  mutate(country = fct_reorder(country, -unemployment)) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = country)) +
  geom_point()

```

Now, create a scatter plot of ``GDP`` against ``research_investment`` in 2014, with a different colour for each point. Limit the number of countries with colours to the **top 5 countries** by GDP:

```{r col1, exercise=TRUE}

```

```{r col1-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  mutate(country = ifelse(row_number() <= 5, country, "Other")) %>% 
  mutate(country = fct_reorder(country, -GDP)) %>% 
  ggplot(aes(x = research_investment, y = GDP, col = country)) +
  geom_point()

```

### Colours for lines

For line charts, colours are a little bit different, as you also have to add a ``group`` variable. For example, this shows GDP over time for France, the UK and Germany:

```{r col_ex6, echo = TRUE}

economic_data %>% 
  filter(country %in% c("France", "UK", "Germany")) %>% 
  ggplot(aes(x = year, y = GDP, group = country, col = country)) +
  geom_line()

```

Now, create a line chart of ``wages`` over time for Italy, Greece and Spain, with a different colour for each country. **Additionally, add ``size = 2`` inside the ``geom_line()`` function to make the lines a bit bigger:

```{r col2, exercise=TRUE}

```

```{r col2-solution}

economic_data %>% 
  filter(country %in% c("Italy", "Greece", "Spain")) %>% 
  ggplot(aes(x = year, y = GDP, group = country, col = country)) +
  geom_line(size = 2)

```


### Colour palettes

Until now, all the colours have been defined automatically by R. But we can change that of course! There is a big range of functions in the family of ``scale_`` that allow you do that.

The following manually create a colour palette of green, red and blue for the previous line chart:

```{r col_ex7, echo = TRUE}

economic_data %>% 
  filter(country %in% c("France", "UK", "Germany")) %>% 
  ggplot(aes(x = year, y = GDP, group = country, col = country)) +
  geom_line(size = 2) +
  scale_colour_manual(values = c("green", "red", "blue"))

```


Copy the previous lines chart you made of Italy, Greece and Spain, and create a manual colour palette of darkblue, lightblue, and red:

```{r col3, exercise=TRUE}

```

```{r col3-solution}

economic_data %>% 
  filter(country %in% c("Italy", "Greece", "Spain")) %>% 
  ggplot(aes(x = year, y = GDP, group = country, col = country)) +
  geom_line(size = 2) +
  scale_colour_manual(values = c("darkblue", "lightblue", "red"))

```


You can also use some ready-made colour palettes which are very helpful, especially from the ``RColorBrewer`` package. You can find a description of the palettes [here](http://www.sthda.com/english/wiki/colors-in-r).

In the example below, we create a scatterplot by country (and "Other") by using the ``Dark2`` palette. This is an appropriate colour palette for categorical data like country names. Note that we also increase the point size to make it more visible:

```{r col_ex8, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(unemployment)) %>% 
  mutate(country = ifelse(row_number() <= 7, country, "Other")) %>% 
  mutate(country = fct_reorder(country, -unemployment)) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = country)) +
  geom_point(size = 4) +
  scale_color_brewer(palette = "Dark2")

```


Now copy the scatterplot above and experiment with a few palettes such as ``Blues``, ``Set3`` or ``Pastel1`` to see the difference:

```{r col4, exercise=TRUE}

```

```{r col4-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(unemployment)) %>% 
  mutate(country = ifelse(row_number() <= 7, country, "Other")) %>% 
  mutate(country = fct_reorder(country, -unemployment)) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = country)) +
  geom_point(size = 4) +
  scale_color_brewer(palette = "Pastel1")

```

We can also create continuous palettes which are useful when the colour represents a numeric variable. In the scatterplot below, for example, colour represents ``wages``:

```{r col_ex9, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(unemployment)) %>% 
  mutate(country = ifelse(row_number() <= 7, country, "Other")) %>% 
  mutate(country = fct_reorder(country, -unemployment)) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = wages)) +
  geom_point(size = 4) +
  scale_color_gradient(low="blue", high="red")

```

Now, you create a similar scatterplot as above, but with the following changes: 

* a color gradient from green to red
* point size proportional to wages
* add ``alpha = 0.5`` inside ``geom_point()`` to add some transparency

```{r col5, exercise=TRUE}

```

```{r col5-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(unemployment)) %>% 
  mutate(country = ifelse(row_number() <= 7, country, "Other")) %>% 
  mutate(country = fct_reorder(country, -unemployment)) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = wages, size = wages)) +
  geom_point(alpha = 0.5) +
  scale_color_gradient(low="green", high="red")

```

## Scale

Sometimes you want to modify the scales for your chart. For example, in the chart below, the y-axis doesn't start at zero. This can be an issue because the chart makes it seem like there is very high growth, but that may not be so clear if the y-axis started at 0:

```{r scale_ex1, echo = TRUE}

economic_data %>% 
  filter(country == "France") %>% 
  ggplot(aes(x = year, y = GDP)) +
  geom_line()

```

Modify the chart to make the y-axis start at zero by adding ``+ expand_limits(y = 0)`` at the end:

```{r scale1, exercise=TRUE}

```

```{r scale1-solution}

economic_data %>% 
  filter(country == "France") %>% 
  ggplot(aes(x = year, y = GDP)) +
  geom_line() +
  expand_limits(y = 0)

```

Notice how the chart is much less steep?

You can also manually change the upper and lower limits of a chart by using ``ylim(min, max)`` and ``xlim(min, max)``:

```{r scale_ex2, echo = TRUE}

economic_data %>% 
  filter(country == "France") %>% 
  ggplot(aes(x = year, y = GDP)) +
  geom_line() +
  xlim(1996, 2010) +
  ylim(0, 3000000)

```


Now, create the following scatterplot:

* a scatterplot of ``wages`` on the y-axis again ``unemployment`` on the x-axis for 2014
* the colour and the transparency (``alpha``) should be proportional to GDP
* the size of all points should be 3
* the x scale should go from 0 to 30
* the y scale should go from 0 to 70000


```{r scale2, exercise=TRUE}

```

```{r scale2-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  ggplot(aes(x = unemployment, y = wages, colour = GDP, alpha = GDP)) +
  geom_point(size = 3) +
  xlim(0, 30) +
  ylim(0, 70000)

```


## Coord_flip

Notice in the chart below, it is difficult to read the labels on the x-axis because of the length of some country names:

```{r scale_ex3, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col()

```


Make it easier to read in that case by flipping the axis and putting the country names on the y-axis instead, by using ``+ coord_flip()`` at the end of the code above:

```{r scale3, exercise=TRUE}

```

```{r scale3-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col() +
  coord_flip()

```


## Labels and title

It is always important to label your chart correctly. Notice in the chart below, the x-axis looks particularly bad because we used the ``reorder()`` function:


```{r labels_ex1, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col() +
  coord_flip()

```

You can add appropriate axis titles, as well as a title for the whole chart using the ``xlab()``, ``ylabl()`` and ``ggtitle()`` functions:

```{r labels_ex2, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col() +
  coord_flip() +
  xlab("Country") +
  ylab("GDP ($millions)") +
  ggtitle("GDP for top 10 largest European economies in 2014")
  

```

Now your turn. Construct the following chart:

* A line chart of govt_debt over time for France, the UK, Germany, Spain, Italy, Greece
* The line size should be 2 for all countries
* There should be a different colour for each country
* The colour scale should use the ``Paired`` palette
* The x-axis title should be "Date (year)"
* The y-axis title should be "Government debt (%)"
* The chart title should be "Government debt over time for selected European countries"

```{r labels1, exercise=TRUE}

```

```{r labels1-solution}

economic_data %>% 
  filter(country %in% c("France", "UK", "Germany", "Spain", "Italy", "Greece")) %>% 
  ggplot(aes(x = year, y = govt_debt, col = country)) +
  geom_line(size = 2) +
  xlab("Date (year)") +
  ylab("Government debt (%)") +
  ggtitle("Government debt over time for selected European countries")

```


## Themes

There are lots of changes you can make to a chart's look - for example:
* the background colour
* the number of grids in the background
* the axis text size
* The title size

These can all be controlled using the **theme()** function in ggplot.

### Text 

Start with this chart:

```{r theme_ex1, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col() +
  xlab("Country") +
  ylab("GDP ($millions)") +
  ggtitle("GDP for top 10 largest European economies in 2014")
  

```

An alternative to flipping the chart could be to make the axis text rotate 45 degrees. We can also try to center the title. Note that with all modifications to text themes, we need to use ``element_text``:


```{r theme_ex2, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col() +
  xlab("Country") +
  ylab("GDP ($millions)") +
  ggtitle("GDP for top 10 largest European economies in 2014") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))
  

```

Now take the chart above and add the following modifications:

* All axis text (``axis.text``) should have size 12, be in bold (``face = "bold"``) and be red (``colour = "red"``)
* The title should have size 20

```{r theme1, exercise=TRUE}

```

```{r theme1-solution}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col() +
  xlab("Country") +
  ylab("GDP ($millions)") +
  ggtitle("GDP for top 10 largest European economies in 2014") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 12, face = "bold", colour = "red"),
        plot.title = element_text(hjust = 0.5, size = 20))

```


### Background

You can also modify the background colour and the colour of the background grids as follows:

```{r theme_ex3, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(GDP)) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(country, -GDP), y = GDP)) +
  geom_col() +
  xlab("Country") +
  ylab("GDP ($millions)") +
  ggtitle("GDP for top 10 largest European economies in 2014") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 12, face = "bold"),
        plot.title = element_text(hjust = 0.5),
        panel.background = element_rect(fill = "red"),
        panel.grid.major = element_line(colour = "grey"))
  

```

### Ready-made themes

In reality, you are moer likely to use ready-made themes which change a number of elements of the chart at a time.

For example, ``theme_bw()`` is one I use a lot, which creates a nice white background with grey grids:

```{r theme_ex4, echo = TRUE}

economic_data %>% 
  filter(year == 2014) %>% 
  arrange(desc(unemployment)) %>% 
  mutate(country = ifelse(row_number() <= 7, country, "Other")) %>% 
  mutate(country = fct_reorder(country, -unemployment)) %>% 
  ggplot(aes(x = unemployment, y = inflation, col = country)) +
  geom_point(size = 4) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
  

```


Now, experiment with the chart above and replace ``theme_bw()``with the following themes to see what they do:

* ``theme_classic()``
* ``theme_dark()``
* ``theme_void()``
* ``theme_economist()``
* ``theme_tufte()``

```{r theme2, exercise=TRUE}

```

```{r theme2-solution}



```


