---
title: |
 | Predicting NFL attendance
 | <font size = 4> American University of Armenia </font>
 | <font size = 4> `r format(Sys.Date(), '%d %b %Y')`
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}

library(tidyverse)
library(tidymodels)
library(ranger)
library(corrplot)
library(vip)

# Set seed and import data

set.seed(52)

source("plot_tuning_metrics.R")


attendance <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/attendance.csv")
standings <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/standings.csv")

attendance_joined <- attendance %>%
  left_join(standings,
            by = c("year", "team_name", "team")
  ) %>%
  filter(!is.na(weekly_attendance)) %>%
  select(
    weekly_attendance, team_name, year, week,
    margin_of_victory, strength_of_schedule, playoffs
  ) %>% 
  mutate(weekly_attendance = ifelse(weekly_attendance > 80000, "high", "other")) %>% 
  mutate(weekly_attendance = as.factor(weekly_attendance),
         team_name = as.factor(team_name),
         playoffs = as.factor(playoffs)) %>% 
  mutate(tech_factor1 = margin_of_victory * 0.5 + rnorm(1, 5),
         tech_factor2 = margin_of_victory * 0.5 + rnorm(1, 10)) 

na_sample <- sample(c(1:nrow(attendance_joined)), 
                    0.1 * nrow(attendance_joined), replace = FALSE)

attendance_joined[na_sample, "strength_of_schedule"] <- NA

```

## Introduction

This document reports the findings of a model used to predict whether an NFL game will have high attendance or not. An NFL game is classified as having high attendance if it is attended by more than 80,000 people.


## Exploratory data analysis

We look at the distribution of the ``margin_of_victory`` variable using the histogram below:

```{r echo = FALSE, warning = FALSE, message = FALSE}

attendance_joined %>% 
  ggplot(aes(x = margin_of_victory)) +
  geom_histogram() +
  theme_bw()
  
```

Additionally, we can see there there is a high degree of correlation between some variables:

```{r echo = FALSE, warning = FALSE, message = FALSE}

attendance_joined %>% 
  na.omit() %>% 
  select(year, week, margin_of_victory, strength_of_schedule, tech_factor1, tech_factor2) %>% 
  cor() %>% 
  corrplot()
  
```


